How Large a Sample Size is Needed for Block Bootstrap Confidence Intervals to Have Desired Coverage Rates?

### 2023-03-08 Code/NewPlots.Rmd
Manuscript Update

```{r}
graph_bts <- function(cov, phi, target, width, nrep = 1000, sizes = c(100, 200, 300, 400, 500), types = c('estCI', 'mbtCI', 'altCI', 'bcaCIest', 'bcaCImbt', 'bcaCIalt'), parameters = c('mu', 'sigma', 'phi'), level = .95) {
  n <- rep(sizes, each = length(types) * length(parameters))
  CI <- rep(types, times = length(sizes), each = length(parameters))
  t <- rep(parameters, times = length(sizes) * length(types))
  LB <- sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[1])
  UB <-sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[2])

  df <- data.frame(n, CI, t, cov, LB, UB)
  library(ggplot2)
  ggplot(df, aes(x = n, y = cov)) +
    geom_hline(yintercept = level, linetype = 'dashed', color = 'orange') + 
    geom_line() +
    geom_errorbar(aes(ymin=LB, ymax = UB), colour="black", width = width) +
    facet_grid(t ~ CI)
}
```

```{r}
phi <- 0
target <- c(0, 1, phi)
cov <- c(mychk(sim_0_1h, target), mychk(sim_0_2h, target), mychk(sim_0_3h, target), mychk(sim_0_4h, target), mychk(sim_0_5h, target))
graph_bts(cov, phi, target, width = 25)
ggsave('../Manuscript/figures/plot_0.pdf')
```
```{r}
phi <- .2
target <- c(0, 1, phi)
cov <- c(mychk(sim_.2_1h, target), mychk(sim_.2_2h, target), mychk(sim_.2_3h, target), mychk(sim_.2_4h, target), mychk(sim_.2_5h, target))
graph_bts(cov, phi, target, width = 25)
ggsave('../Manuscript/figures/plot_.2.pdf')
```
```{r}
phi <- .4
target <- c(0, 1, phi)
cov <- c(mychk(sim_.4_1h, target), mychk(sim_.4_2h, target), mychk(sim_.4_3h, target), mychk(sim_.4_4h, target), mychk(sim_.4_5h, target))

graph_bts(cov, phi, target, width = 10)
ggsave('../Manuscript/figures/plot_.4.pdf')
```

>>>>

```{r}
df_bts <- function(cov, nrep = 1000, phis = c(-0.4, -0.2, 0, 0.2, 0.4), sizes = c(100, 200, 300, 400, 500, 600, 700), types = c('pctCI', 'estCI', 'bcaCIest'), parameters = c('mu', 'sigma', 'phi')) {
  phi <- rep(phis, each = length(sizes) * length(types) * length(parameters))
  n <- rep(sizes, times = length(phis), each = length(types) * length(parameters))
  CI <- rep(types, times = length(phis) * length(sizes), each = length(parameters))
  t <- rep(parameters, times = length(phis) * length(types) * length(sizes))
  LB <- sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[1])
  UB <- sapply(cov, function(i) prop.test(i*nrep, nrep)$conf.int[2])

  c(data.frame(phi, n, CI, t, cov, LB, UB))
}

```

```{r}
target <- c(0, 1, -.4)
cov <- c(mychk(sim_n.4_1h, target), mychk(sim_n.4_2h, target), mychk(sim_n.4_3h, target), mychk(sim_n.4_4h, target), mychk(sim_n.4_5h, target), mychk(sim_n.4_6h, target), mychk(sim_n.4_7h, target))
target <- c(0, 1, -.2)
cov <- c(cov, mychk(sim_n.2_1h, target), mychk(sim_n.2_2h, target), mychk(sim_n.2_3h, target), mychk(sim_n.2_4h, target), mychk(sim_n.2_5h, target), mychk(sim_n.2_6h, target), mychk(sim_n.2_7h, target))
target <- c(0, 1, 0)
cov <- c(cov, mychk(sim_0_1h, target), mychk(sim_0_2h, target), mychk(sim_0_3h, target), mychk(sim_0_4h, target), mychk(sim_0_5h, target), mychk(sim_0_6h, target), mychk(sim_0_7h, target))
target <- c(0, 1, .2)
cov <- c(cov, mychk(sim_.2_1h, target), mychk(sim_.2_2h, target), mychk(sim_.2_3h, target), mychk(sim_.2_4h, target), mychk(sim_.2_5h, target), mychk(sim_.2_6h, target), mychk(sim_.2_7h, target)) 
target <- c(0, 1, .4)
cov <- c(cov, mychk(sim_.4_1h, target), mychk(sim_.4_2h, target), mychk(sim_.4_3h, target), mychk(sim_.4_4h, target), mychk(sim_.4_5h, target), mychk(sim_.4_6h, target), mychk(sim_.4_7h, target))

df <- data.frame(df_bts(cov))
graph_bts <- function(t, width, data, trans = 'identity', level = .95)
{
  library(latex2exp)
  library(ggplot2)

  ggplot(data, aes(x = n, y = cov)) +
    geom_hline(yintercept = level, linetype = 'dashed', color = 'orange') + 
    geom_line() +
    geom_errorbar(aes(ymin=LB, ymax = UB), colour="black", width = width) +
    facet_grid(factor(phi) ~ factor(CI), labeller = label_parsed) +
    scale_y_continuous(trans=trans)
  ggsave(paste('../Manuscript/figures/plot_', t, '.pdf', sep = ''))
}
```

```{r}
t <- 'mu'
width <- 10
df1 <- df[(df$t == t),]
graph_bts(t, width, df1)
```

```{r}
t <- 'sigma'
width <- 10
df2 <- df[(df$t == t),]
graph_bts(t, width, df2)
```

```{r}
t <- 'phi'
width <- 10
df3 <- df[(df$t == t),]
graph_bts(t, width, df3)
```

### 2023-03-08 Manuscript/manuscript.tex
Manuscript Update

\title{How Large a Sample Size is Needed for Block Bootstrap Confidence Intrvals
  to Have Desired Coverage Rates?}

@@

The goal of this paper is to provide some practical recommendations on necessary
sample size for block bootstrap with dependent data. We consider a simple
situation of a stationary time series, where the parameters of interests are the
mean, standard deviation, and the first-order autocorrelation coefficient.
We compare four variants of confidence intervals from block bootstrap:
\jy{list them} literature \citep{diciccio1996bootstrap}
\citep{rice2006mathematical}.
Their empirical coverage rates at different sample sizes and dependence levels
are compared in a simulation study.
\jy{may reveal the most interesting result here to engage interests.}

The remainder of the papers is organized as follows.
\jy{refer to sections by label instead of hard number}
these intervals are constructed as well as the general procedure for block
bootstrap parameter estimation is reviewed in Section 2. Section 3 offers an
explanation of the simulation design and results. Section 4 discusses the main
takeaways from the results, how these findings could be useful, and possible
future explorations.

@@

In basic bootstrap procedure, the new pseudo-sample of size $n$ would be
created by simply resampling observations from the original sample with
replacement. However, in the case of a time series, in order to account for
the temporal dependence, the time series can be split into blocks, typically
of the same size. The block should be of size $l$ large enough for each
pseudo-sample to exhibit some temporal dependence, yet small enough for there
to be some variability between each pseudo-sample. Ideally, as $n$ increases,
$l$ should also increase, but $l / n$ should decrease. To achieve this, $l$ is
often assigned a value as a function of $n$. A common function that is
considered the best by much previous literature is $l = \lceil n^{1/3} \rceil$
\citep{buhlmann1999block}. The designer of the study can choose whether the
blocks overlap or not. If the blocks do not overlap, it is called a non-moving
block bootstrap. In this case, the original sample is split evenly into
blocks, which will then be resampled to create new pseudo-samples. If the
blocks do overlap, it is called a moving block bootstrap. Blocks are resampled
randomly from the original sample, but they do not have to be from a set of
evenly spaced blocks. For moving block bootstrap, if $l \vert n$, a
bootstrapped pseudo-sample is created by taking a sample of $n / l$ blocks of
size $l$ with replacement to create a new pseudo-sample of size $n$. If
$l does not divide n$, the pseudo-sample is created by taking a sample of
$\lfloor n / l \rfloor$ blocks of size $l$ and one block of size
$(n \;\mathrm{mod})\; l$. An estimate $\hat\theta_n^*$ is computed from the
pseudo-sample. This procedure can be repeated many times to create a
distribution of $\hat\theta_n^*$.

In this study, 1000 $\hat\theta_n^*$ were created for one simulation of block
bootstrap estimation. Using this distribution of $\hat\theta_n^*$, a
$1 - \alpha$/2 \% confidence interval for the parameter can be constructed. 
The most simplistic method for constructing this interval is the percentile
confidence interval. For a parameter $\theta$, the percentile $1 - \alpha$/2
\% confidence interval takes the form: 
\[ [\hat\theta_{n, \alpha/2}^*, \hat\theta_{n, 1 - \alpha/2}^*].\] 
Another method is the bias-corrected and accelerated 
(BCA) $1 - \alpha$/2 \% confidence interval, which takes the form: 
\[ [\hat\theta_{n, \alpha_1}^*,\hat\theta_{n, \alpha_2}^*].\] 
$\alpha_{1}$ and $\alpha_{2}$ are the cumulative probability of $z_{1}$ and
$z_{2}$, respectively, where:
\[z_{1} = \frac{z_{0} - z_{1 - \alpha/2}}{1 - a(z_{0} - z_{1 - \alpha/2})} +
z_{0}\] and
\[z_{2} = \frac{z_{0} + z_{1 - \alpha/2}}{1 - a(z_{0} + z_{1 - \alpha/2})} +
z_{0}.\] 
Where $z_0$ is the quantile function of the proportion of
$\hat\theta_n^* < \bar\theta_n^*$, and $a$ is the skewness of
$\hat{\theta}_{n, -i}$, where $\hat{\theta}_{n, -i}$ is the statistic of the
original sample computed without the $i^{th}$ block.

Both of these intervals work well when attempting to recover the mean and
standard deviation of a temporally dependent process, but their coverage of
the temporal dependence deteriorates as $n$ increases. A potential solution to
this issue is to center the intervals $\hat{\theta}_{n}$. To do this, the
distribution $\delta^*$ should be considered. A percentile interval centered
around $\hat{\theta}_{n}$ ($CI_{\hat{\theta}_{n}}$) takes the form:
\[ [\hat{\theta}_{n} + \delta^*_{\alpha/2},
  \hat{\theta}_{n} + \delta^*_{1 - \alpha/2}].\] 
This centering procedure can also be applied to a BCA interval. A BCA interval
centered around $\hat{\theta}_{n}$ ($CI_{BCA, \hat{\theta}_{n}}$) takes the
form:
\[ [\hat{\theta}_{n} + \delta^*_{\alpha_1},
  \hat{\theta}_{n} + \delta^*_{\alpha_2}].\]

Thus, there are two types of intervals that can be constructed for block
bootstrap estimation in this study: $CI_{\hat{\theta}_{n}}$ and
$CI_{BCA, \hat{\theta}_{n}}$. Note that both of these intervals are centered
around $\hat{\theta}_{n}$ and asymmetric.

@@

Base R has a built in function arima.sim that allows one to simulate an AR(1)
process, for which one can set the theoretical $\phi$. By default, the mean of
the process $\mu$ is 0 and the standard deviation of the error term
$\sigma_{\epsilon}$ is 1. For the purposes of this study, the standard
deviation of the process $\sigma_{x}$ should be controlled at 1. The
observations can be multiplied by $\sqrt{1 - \phi^2}$ so that the new
theoretical $\sigma_{x}$ is equivalent to 1, as shown below:
\begin{align}
X_{t} &= \epsilon_{t} + \phi X_{t-1};\\
Var(X_{t}) &= Var(\epsilon_{t} + \phi X_{t-1});\\
\sigma^2_{x} &= \sigma^2_{\epsilon} + \phi^2 \sigma^2_{x};\\
\sigma^2_{x}(1 - \phi^2) &= 1;\\
\sigma_{x}\sqrt{1 - \phi^2} &= 1.
\end{align}
The experimental factor of this study was $\phi$, as the temporal dependence
affects the performance of block bootstrap.
$\phi \in \{-0.4 -0.2, 0, 0.2, 0.4\}$.

As described in the last section, the block bootstrap method can be used to
estimate a parameter $\theta$ of an AR(1) process. This method can be
simulated with the tsboot function from the boot R package. In this study,
$\theta$ is composed of the target parameters $\mu = 0$, $\sigma_{x}$, and
$\phi$. The sample mean $\bar{x}$ can be used as the statistic to estimate
$\mu$, the sample standard deviation $S_x$ can be used as the statistic to
estimate $\sigma$, and the autocorrelation function of the sample at lag 1
$\rho_1$ can be used as the statistic to estimate $\phi$. In the R
implementation, these three statistics are incorporated into a function that
will apply each statistic to the original sample and to each bootstrapped
pseudo-sample. From a sample of length $n$, two types of block bootstrap 95\%
confidence intervals can be constructed for each target parameter:
$CI_{\hat{\theta}_{n}}$ and $CI_{BCA, \hat{\theta}_{n}}$.

@@

\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_n.4}
  \caption{At $\phi = -.4$, both intervals covered $\mu$ correctly as low as
    $n = 100$. $CI_{\hat{\theta}_{n}}$ seems to cover $\sigma$ well as low as
    $n = 400$. $CI_{BCA, \hat{\theta}_{n}}$ under-covers $\sigma$, but there
    is no clear pattern of deterioration of coverage, so one can expect that
    increasing $n$ will eventually result in proper coverage of $\sigma$.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration, and this
    is true as low as $n = 200$. $CI_{BCA, \hat{\theta}_{n}}$ greatly
    under-covers $\phi$, and there is a clear pattern of deterioration of
    coverage.}
  \label{fig:plot_n.4}
\end{figure}

>>>>

\title{How Large a Sample Size is Needed for Block Bootstrap Confidence
  Intervals
  to Have Desired Coverage Rates?}

@@

The goal of this paper is to provide some practical recommendations on
necessary sample size for block bootstrap with dependent data. We consider a
simple situation of a stationary time series, where the parameters of
interests are the mean, standard deviation, and the first-order
autocorrelation coefficient. We compare three variants of block bootstrap
confidence intervals from the literature \citep{diciccio1996bootstrap}
\citep{rice2006mathematical}: a basic percentile interval, a percentile
interval centered around the original estimate, and a bias-corrected and
accelerated ($BC_a$) interval also centered around the original estimate. Their
empirical coverage rates at different sample sizes and dependence levels are
compared in a simulation study. The results of this study suggest that
appropriate recovery of temporal dependence parameters is heavily dependent on
the type of interval used.

The remainder of the papers is organized as follows:
The Block Bootstrap Confidence Intervals section reviews how these intervals
are constructed as well as the general procedure for block bootstrap parameter
estimation. The Simulation Study section offers an explanation of the
simulation design and results. Finally, the Concluding Remarks section
discusses the main takeaways from the results, how these findings could be
useful, and possible future explorations.

@@

In basic bootstrap procedure, the new pseudo-sample of size $n$ would be
created by simply resampling observations from the original sample with
replacement. However, in the case of a time series, in order to account for
the temporal dependence, the time series can be split into blocks, typically
of the same size. The block should be of size $l$ large enough for each
pseudo-sample to exhibit some temporal dependence, yet small enough for there
to be some variability between each pseudo-sample. Ideally, as $n$ increases,
$l$ should also increase, but $l / n$ should decrease. To achieve this, $l$ is
often assigned a value as a function of $n$. A common function that is
considered the best by much previous literature is $l = \lceil n^{1/3} \rceil$
\citep{buhlmann1999block}.

The designer of the study can choose whether the blocks overlap or not. If the
blocks do not overlap, it is called non-overlapping or non-moving block
bootstrap. In this case, blocks are resample randomly from the set of all
non-overlapping blocks of size $l$. For non-moving block bootstrap, $l \vert n$
must always be true. In this case, a set of $n / l$ non-overlapping blocks of
size $l$ is constructed from the original sample. Each bootstrapped
pseudo-sample is created by sampling $n / l$ blocks with replacement from the
same set.

If the blocks do overlap, it is called overlapping or moving block bootstrap.
In this case, blocks are resampled randomly from the set of all possible
blocks of size $l$. For moving block bootstrap, $l \vert n$ is not a necessary
condition. The set of available blocks is created by iterating through each
observation $i$ in the original sample and adding to the set the block of size
$l$ that begins with $i$. This set can limited to blocks that end with an
$i <= n$, meaning that $n - l + 1$ blocks are availabe to resample from, or it
can be expanded to include blocks that ``wrap around'': in
other words, blocks that end with a lesser index than the index they began
with, meaning that $n$ blocks are available to resample from. If $l \vert n$,
a bootstrapped pseudo-sample is created by taking a sample of $n / l$ blocks
from the set with replacement to create a new pseudo-sample of size $n$. If
$l does not divide n$, the last block selected will be reduced in size so that
the final size of the pseudo-sample is $n$. Moving block bootstrap was used in
this study as it has been found to yield the lowest error in general
\citep{radovanov2014comparison}. The set of available was not limited as
described above.

An estimate $\hat\theta_n^*$ is computed from the pseudo-sample. This
procedure can be repeated many times to create a distribution of
$\hat\theta_n^*$.

In this study, 1000 $\hat\theta_n^*$ were created for one simulation of block
bootstrap estimation. Using this distribution of $\hat\theta_n^*$, several
variants of a $1 - \alpha$/2 \% confidence interval for the parameter can be
constructed. The most simplistic method for constructing this interval is the
basic percentile confidence interval. For a parameter $\theta$, the percentile
$1 - \alpha$/2 \% confidence interval takes the form: 
\[ [\hat\theta_{n, \alpha/2}^*, \hat\theta_{n, 1 - \alpha/2}^*].\]

This interval works well when attempting to recover the mean and
standard deviation of a temporally dependent process, but its coverage of
the temporal dependence deteriorates as $n$ increases. A potential solution to
this issue is to center the intervals around $\hat{\theta}_{n}$. To do this,
the distribution of $\delta^*$ should be considered. For a parameter $\theta$,
the $\hat{\theta}_{n}$-centered $1 - \alpha$/2 \% confidence interval takes
the form:
\[ [\hat{\theta}_{n} + \delta^*_{\alpha/2},
  \hat{\theta}_{n} + \delta^*_{1 - \alpha/2}].\]

An alternative method is a $BC_a$ interval centered around $\hat{\theta}_{n}$.
For a parameter $\theta$, the $BC_a$ $1 - \alpha$/2 \% confidence interval
takes the form:
\[ [\hat{\theta}_{n} + \delta^*_{\alpha_1},
  \hat{\theta}_{n} + \delta^*_{\alpha_2}].\] 
$\alpha_{1}$ and $\alpha_{2}$ are the cumulative probability of $z_{1}$ and
$z_{2}$, respectively, where:
\[z_{1} = \frac{z_{0} - z_{1 - \alpha/2}}{1 - a(z_{0} - z_{1 - \alpha/2})} +
z_{0}\] and
\[z_{2} = \frac{z_{0} + z_{1 - \alpha/2}}{1 - a(z_{0} + z_{1 - \alpha/2})} +
z_{0}.\] 
Where $z_0$ is the quantile function of the proportion of
$\hat\theta_n^* < \bar\theta_n^*$, and $a$ is the skewness (computed using
the skewness function from the R e1071 package \citep{e1071}) of
$\hat{\theta}_{n, -i}$, where $\hat{\theta}_{n, -i}$ is the statistic of the
original sample computed without the $i^{th}$ block.

Thus, there are three types of intervals that can be constructed for block
bootstrap estimation in this study: Percentile, $\hat{\theta}_{n}$-Centered,
and $BC_a$.

@@

The Base R stats package has a built in function arima.sim that allows one to
simulate an AR(1) process, for which one can set the theoretical $\phi$
\citep{R}. By default, the mean of the process $\mu$ is 0 and the standard
deviation of the error term $\sigma_{\epsilon}$ is 1. For the purposes of this
study, the standard deviation of the process $\sigma_{x}$ should be controlled
at 1. The observations can be multiplied by $\sqrt{1 - \phi^2}$ so that the new
theoretical $\sigma_{x}$ is equivalent to 1, as shown below:
\begin{align}
X_{t} &= \epsilon_{t} + \phi X_{t-1};\\
Var(X_{t}) &= Var(\epsilon_{t} + \phi X_{t-1});\\
\sigma^2_{x} &= \sigma^2_{\epsilon} + \phi^2 \sigma^2_{x};\\
\sigma^2_{x}(1 - \phi^2) &= 1;\\
\sigma_{x}\sqrt{1 - \phi^2} &= 1.
\end{align}
The experimental factor of this study was $\phi$, as the temporal dependence
affects the performance of block bootstrap.
$\phi \in \{-0.4 -0.2, 0, 0.2, 0.4\}$.

As described in the last section, the block bootstrap method can be used to
estimate a parameter $\theta$ of an AR(1) process. This method can be
simulated with the tsboot function from the R boot package \citep{boot}
\citep{davison}. In this study, $\theta$ is composed of the target parameters
$\mu = 0$, $\sigma_{x}$, and $\phi$. The sample mean $\bar{x}$ can be used as
the statistic to estimate $\mu$, the sample standard deviation $S_x$ can be
used as the statistic to estimate $\sigma$, and the autocorrelation function
of the sample at lag 1 $\rho_1$ can be used as the statistic to estimate
$\phi$. In the R implementation, these three statistics are incorporated into
a function that will apply each statistic to the original sample and to each
bootstrapped pseudo-sample. From a sample of length $n$, two types of block
bootstrap 95\% confidence intervals can be constructed for each target
parameter:
$CI_{\hat{\theta}_{n}}$ and $CI_{BCA, \hat{\theta}_{n}}$.

@@

The figures below were created using the R ggplot2 package \citep{ggplot2}.
\begin{figure}[tbp]
  \centering
  \includegraphics[width=\textwidth]{figures/plot_n.4}
  \caption{At $\phi = -.4$, both intervals covered $\mu$ correctly as low as
    $n = 100$. $CI_{\hat{\theta}_{n}}$ seems to cover $\sigma$ well as low as
    $n = 400$. $CI_{BCA, \hat{\theta}_{n}}$ under-covers $\sigma$, but there
    is no clear pattern of deterioration of coverage, so one can expect that
    increasing $n$ will eventually result in proper coverage of $\sigma$.
    $CI_{\hat{\theta}_{n}}$ covers $\phi$ well with no deterioration, and this
    is true as low as $n = 200$. $CI_{BCA, \hat{\theta}_{n}}$ greatly
    under-covers $\phi$, and there is a clear pattern of deterioration of
    coverage.}
  \label{fig:plot_n.4}
\end{figure}
